---
title: "EGRET and rloadest Tutorial"
author: "<b>USGS</b>: Laura DeCicco"
date: "`r format(Sys.time(), '%d %B, %Y')`"
knit: (function(inputFile, encoding) { 
      out_dir <- 'test';
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_file='../EGRET.html') })
output:
  ioslides_presentation:
    css: ../styles.css
    logo: ../images/simple-shadow.png
    smaller: yes
    widescreen: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

What is `EGRET`?

Exploration and Graphics for RivEr Trends (EGRET)

EGRET includes statistics and graphics for streamflow history, water quality trends, and the statistical modeling algorithm Weighted Regressions on Time, Discharge, and Season (WRTDS).

A very complete document can be found [HERE](http://pubs.usgs.gov/tm/04/a10/)

## EGRET Installation

`EGRET` is available on the CRAN repository. The CRAN version is the most stable and user-tested:

```{r echo=TRUE, eval=FALSE}
install.packages("EGRET")
```

Bug fixes and feature upgrades are vetted through a version of `EGRET` that is available on a USGS-maintained R repository. To install from that repository:

```{r echo=TRUE, eval=FALSE}
install.packages("EGRET", 
                 repos=c("http://owi.usgs.gov/R",
                         getOption("repos")))
```

More information can be found [here](http://owi.usgs.gov/R/gran.html).

Finally, the absolute cutting-edge version of `EGRET` can be installed using the `devtools` package which pulls from GitHub:

```{r echo=TRUE, eval=FALSE}
library(devtools)
install_github("USGS-R/EGRET")
```

## EGRET Help

Once the `EGRET` package has been installed, it needs to be loaded in order to use any of the functions:

```{r echo=TRUE, eval=TRUE}
library(EGRET)
```

There is a vignette that covers the full scope of the `EGRET` package. It can be accessed with the following command:

```{r echo=TRUE, eval=FALSE}
vignette("EGRET",package = "EGRET")
```

Additionally, each function has a help file. These can be accessed by typing a question mark, followed by the function name in the R console:

```{r echo=TRUE, eval=FALSE}
?modelEstimation
```

Finally, if there are still questions that the vignette and help files don't answer, please post an issue on the `EGRET` GitHub page:

<center>[https://github.com/USGS-R/EGRET/issues](https://github.com/USGS-R/EGRET/issues)</center>

## Yahara Example

```{r echo=TRUE, eval=FALSE}
library(EGRET)

site <- "05427718"
pCode <- "00665"

Sample <- readNWISSample(site, pCode, "1990-01-01","2016-01-01")
Daily <- readNWISDaily(site,"00060", "1990-01-01","2016-01-01")
INFO <- readNWISInfo(site,pCode,interactive = FALSE)

eList <- mergeReport(INFO,Daily,Sample)
eList <- modelEstimation(eList)

saveRDS(eList, "eList_Yahara.rds")

```


Let's explore [HERE](http://waterdatascience.shinyapps.io/exploreEGRET)

## rloadest Installation

`rloadest` is NOT available on the CRAN repository. Therefore, the most stable and user-tested version is on the Geological Survey R Archive Network (GRAN). To install from that repository:

```{r echo=TRUE, eval=FALSE}
install.packages("rloadest", 
                 repos=c("http://owi.usgs.gov/R",
                         getOption("repos")))
```

More information can be found [here](http://owi.usgs.gov/R/gran.html).

Finally, the absolute cutting-edge version of `rloadest` can be installed using the `devtools` package which pulls from GitHub:

```{r echo=TRUE, eval=FALSE}
library(devtools)
install_github("USGS-R/rloadest")
```

## rloadest Help

Once the `rloadest` package has been installed, it needs to be loaded in order to use any of the functions:

```{r echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
library(rloadest)
```

There are several vignettes that covers the full scope of the `rloadest` package. They can be accessed with the following commands:

```{r echo=TRUE, eval=FALSE}
vignette("app1",package = "rloadest")
vignette("app2",package = "rloadest")
vignette("app3",package = "rloadest")
vignette("app4",package = "rloadest")
vignette("app5",package = "rloadest")
vignette("app6",package = "rloadest")
vignette("IncorporatingHysteresis",package = "rloadest")
vignette("InstantaneousTimeStep",package = "rloadest")
vignette("SubDailyTimeSteps",package = "rloadest")
vignette("UsingEGRETData",package = "rloadest")
```


Finally, if there are still questions that the vignette and help files don't answer, please post an issue on the `rloadest` GitHub page:

<center>[https://github.com/USGS-R/rloadest/issues](https://github.com/USGS-R/rloadest/issues)</center>


## rloadest

```{r echo=TRUE, eval=FALSE}
library(rloadest)
library(survival)

Sample <- eList$Sample
Sample <- Sample[!duplicated(Sample$Date),]

loadestModel <- loadReg(Surv(ConcLow, ConcHigh, type="interval2")
                        ~ model(9), 
                   data = Sample, 
                   flow = "Q", dates = "Date",
                   flow.units="cms",
                   conc.units="mg/l",
                   station=eList$INFO$station.nm)

```


## Convert rloadest to EGRET output:

```{r echo=TRUE, eval=FALSE}
Daily <- eList$Daily
Sample <- eList$Sample

# Make DailyLoadest
DailyLoadest <- Daily[,which(!(names(Daily) %in% 
      c("ConcDay","FluxDay","FNConc","FNFlux","SE","yHat")))]

concs <- predConc(loadestModel,DailyLoadest, by="day")
flux <- predLoad(loadestModel,DailyLoadest, by="day")

predictResp <- fitted(loadestModel$cfit, type='response')
predictResp_mean <- fitted(loadestModel$cfit, type='mean')

DailyLoadest$ConcDay <- concs$Conc
DailyLoadest$FluxDay <- DailyLoadest$ConcDay*86.4*DailyLoadest$Q
DailyLoadest$SE <- concs$Std.Err

# Make SampleLoadest

SampleLoadest <- Sample[,which(!(names(Sample) %in% 
        c("yHat","SE","ConcHat")))]

SampleLoadest$SE <- concs$Std.Err[
  which(concs$Date %in% SampleLoadest$Date)]
SampleLoadest$yHat <-predictResp
SampleLoadest$ConcHat <- predictResp_mean

eListLoadest <- as.egret(INFO, DailyLoadest, SampleLoadest)

```


## Pre-defined models

Here are the other pre-defined models that can be used:

Left   Center  
------ ----------
1      $\ln(Load)= \alpha_0 + \alpha_1{\ln{Q}} + \epsilon$
2      $\ln(Load)= \alpha_0 + \alpha_1{\ln{Q}} + \alpha_2{\ln{Q^2}}+ \epsilon$
3      $\ln(Load)= \alpha_0 + \alpha_1{\ln{Q}} + \alpha_2{DecTime}+ \epsilon$
4      $\ln(Load)= \alpha_0 + \alpha_1{\ln{Q}} + \alpha_2{sin(2*\pi{DecTime})} + \alpha_3 cos(2*\pi{DecTime}) + \epsilon$
5      $\ln(Load)= \alpha_0+\alpha_1{\ln{Q}} + \alpha_2{\ln{Q^2}}+\alpha_3{DecTime}+\epsilon$
6      $\ln(Load)=\alpha_0 + \alpha_1{\ln{Q}}+\alpha_2{\ln{Q^2}}+\alpha_3 sin(2\pi{DecTime})+\alpha_4 cos(2\pi{DecTime})+\epsilon$
7      $\ln(Load)=\alpha_0+\alpha_1{\ln{Q}}+\alpha_2 \sin(2\pi{DecTime})+\alpha_3 \cos(2\pi{DecTime})+\alpha_4{DecTime}+\epsilon$ 
8      $\ln(Load)=\alpha_0+\alpha_1{\ln{Q}}+\alpha_2{\ln{Q^2}}+\alpha_3 \sin(2\pi{DecTime})+\alpha_4 \cos(2\pi{DecTime})+\alpha_5{DecTime}+\epsilon$
9      $\ln(Load)=\alpha_0+\alpha_1{\ln{Q}}+\alpha_2{\ln{Q^2}}+\alpha_3 \sin(2\pi{DecTime})+\alpha_4 \cos(2\pi{DecTime})+\alpha_5{DecTime}+\alpha_6{DecTime}^2+\epsilon$   

Where $ln{Q}$ and `DecTime`  are centered using the equations:
$\ln{Q}=\ln(Q)-\mbox{center of } \ln(Q)$
$DecTime=DecYear-\mbox{center of } DecYear$
